---
title: "Analysis GDM PSO&GAs"
author: "Miguel Benito Parejo"
date: "2024-05-04"
output: html_document
---

```{r}
library(readxl)
library(dplyr)
library(ggplot2)
library(car)
library(nortest)
library(jsonlite)
```

# PSO

```{r loading data}
#PSOs6 <- read_excel("files/Results/PSOs.xlsx")
#Gen6_5 <- read_excel("files/Results/GAs/6 5.xlsx")

PSOs <- read_excel("files/Results/PSOs.xlsx")
directory_path <- "files/Results/GAs"

files <- list.files(directory_path, pattern = "\\.xlsx$", full.names = TRUE);

data_list <- lapply(files, function(file) {
  data <- read_excel(file)
  return(data)
})

Gens <- bind_rows(data_list)

```

## Hypervolume study

We start studying the homoscedasticity and normality of the HVs for each configuration.

```{r}
# First we set an ID for each configuration

Master_Configurations <- PSOs %>% 
  select(Pop_Size, Num_Iters, Experts, 
         Alternatives, D_Experts, D_Consensus, Initialisation, Best_Dir) %>% 
  distinct()

Master_Configurations$ID <- 1:nrow(Master_Configurations)
PSOs <- merge(PSOs, Master_Configurations, 
               by = c("Pop_Size", "Num_Iters", "Experts", 
         "Alternatives", "D_Experts", "D_Consensus", "Initialisation",
         "Best_Dir"))

#######

Master_Problem_Type <- PSOs %>% 
  select(Experts, Alternatives, D_Experts, D_Consensus) %>% 
  distinct()

Master_Problem_Type$ID_Type_P <- 1:nrow(Master_Problem_Type)

PSOs <- merge(PSOs, Master_Problem_Type, 
               by = c("Experts", "Alternatives", "D_Experts", "D_Consensus"))

Gens <- merge(Gens, Master_Problem_Type,
                by = c("Experts", "Alternatives", "D_Experts", "D_Consensus"))


Master_PSO_Configuration_Type <- PSOs %>% 
  select(Pop_Size, Num_Iters, Best_Dir, Initialisation) %>% 
  distinct()

Master_PSO_Configuration_Type$ID_Type_C <-
      1:nrow(Master_PSO_Configuration_Type)

Master_GA_Configuration_Type <- Gens %>% 
  select(Pop_Size, Num_Iters, Initialisation, Selection, 
         Crossover, Mutation, Replacement) %>% 
  distinct()

Master_GA_Configuration_Type$ID_Type_C <- 
  nrow(Master_PSO_Configuration_Type)+(1:nrow(Master_GA_Configuration_Type))


PSOs <- merge(PSOs, Master_PSO_Configuration_Type, 
               by = c("Pop_Size", "Num_Iters", "Best_Dir", "Initialisation"))

Gens <- merge(Gens, Master_GA_Configuration_Type, 
               by = c("Pop_Size", "Num_Iters", "Initialisation", "Selection", 
         "Crossover", "Mutation", "Replacement"))

rm(Master_GA_Configuration_Type)
rm(Master_PSO_Configuration_Type)
rm(Master_Configurations)
rm(Master_Problem_Type)
```

```{r}
# Points to number

for(i in 1:50){
  PSOs[sprintf("Point_%s_Consensus",i)] <- pull(PSOs, sprintf("Point_%s",i)) %>% 
    sapply(function(s){
      if(is.na(s)){
        NA
      } else {
        strsplit(s, ",")[[1]][1] %>% 
          gsub('\\{"consensus":', "", .) %>% 
          as.numeric()
      }
    })
  
  PSOs[sprintf("Point_%s_Distance",i)] <- pull(PSOs, sprintf("Point_%s", i)) %>% 
    sapply(function(s){
      if(is.na(s)){
        NA
      } else {
        strsplit(s, ",")[[1]][2] %>% 
          gsub('"distance":', "", .) %>% 
          gsub("\\}", "", .) %>% 
          as.numeric()
      }
    })
}

# Removing texts
PSOs <- PSOs %>% select(-sprintf("Point_%s", 1:50))
```

```{r}
# Points to number

for(i in 1:50){
  Gens[sprintf("Point_%s_Consensus",i)] <- pull(Gens, sprintf("Point_%s",i)) %>% 
    sapply(function(s){
      if(is.na(s)){
        NA
      } else {
        strsplit(s, ",")[[1]][1] %>% 
          gsub('\\{"consensus":', "", .) %>% 
          as.numeric()
      }
    })
  
  Gens[sprintf("Point_%s_Distance",i)] <- pull(Gens, sprintf("Point_%s", i)) %>% 
    sapply(function(s){
      if(is.na(s)){
        NA
      } else {
        strsplit(s, ",")[[1]][2] %>% 
          gsub('"distance":', "", .) %>% 
          gsub("\\}", "", .) %>% 
          as.numeric()
      }
    })
}

# Removing texts
Gens <- Gens %>% select(-sprintf("Point_%s", 1:50))
```


## Juntamos los dataframes

```{r}
# Dataframe joining: Problem type, ID_Configuration, HV and Epsilon

PSO_Probs <- PSOs %>% 
  select(Experts, Alternatives, D_Consensus, D_Experts, 
         ID_Type_C, ID_Type_P, HV, Epsilon)

GEN_Probs <- Gens %>% 
  select(Experts, Alternatives, D_Consensus, D_Experts, 
         ID_Type_C, ID_Type_P, HV, Epsilon)

Problems <- union_all(PSO_Probs, GEN_Probs)
```

```{r}
Problem_ID <- NULL
P_Value <- NULL

for(i in unique(Problems$ID_Type_P)){
  
  #print(sprintf("Fligner-Killen Test: Problem %s", i))
  aux <- fligner.test(Epsilon ~ as.factor(ID_Type_C), 
                     data = Problems %>% filter(ID_Type_P == i))
  #print(aux)
  
  Problem_ID <- c(Problem_ID, i)
  P_Value <- c(P_Value, aux$p.value)
  
}

Result <- data.frame(Problem_ID, P_Value)

```

Due to the lack of homoscedasticity in HV, for each of the problems, conditioned by the configurations, we proceed to perform the Kruskal-Wallis test to compare statistical supremacy among the configurations for each problem.

```{r}
Problem_ID <- NULL
P_Value <- NULL

for(i in unique(Problems$ID_Type_P)){
  
  #print(sprintf("Kruskal-Wallis Test: Problem %s", i))
  aux <- kruskal.test(Epsilon ~ as.factor(ID_Type_C), 
                     data = Problems %>% filter(ID_Type_P == i))
  #print(aux)
  
  Problem_ID <- c(Problem_ID, i)
  P_Value <- c(P_Value, aux$p.value)
  
}

Result_Kruskal_Eps <- data.frame(Problem_ID, P_Value)
```

KW test tells us there are differences for each of the problems conditioned by the configuration.

```{r}
Problem_ID <- NULL
P_Value <- NULL

pairwise_results <- list()

for(i in unique(Problems$ID_Type_P)){
  
#  print(sprintf("Pairwise Wilcox Test: Problema %s", i))
  aux <- pairwise.wilcox.test(
    Problems %>% filter(ID_Type_P == i) %>% pull(HV), 
    Problems %>% filter(ID_Type_P == i) %>% pull(ID_Type_C),
    alternative = "greater",
    exact = F)
 # print(aux)
  
#  Problem_ID <- c(Problem_ID, i)
#  P_Value <- c(P_Value, aux$p.value)
   pairwise_results[[as.character(i)]] <- aux$p.value 
 
}

#Result_Wilcox_HV <- data.frame(Problem_ID, P_Value)



all_ids <- unique(unlist(lapply(pairwise_results, colnames)))
n <- length(all_ids)
p_value_matrix <- matrix(NA, n, n, dimnames = list(all_ids, all_ids))


valid_columns <- which(apply(pairwise_results[[1]], 2, function(col) all(is.na(col) | col >= 0.05)))
print(valid_columns)
print(pairwise_results[[3]][,1])
for(i in names(pairwise_results)) {
#  p_value_matrix[rownames(pairwise_results[[i]]), colnames(pairwise_results[[i]])] <- pairwise_results[[i]]
valid_columns <- which(apply(pairwise_results[[i]], 2, function(col) all(is.na(col) | col >= 0.05)))
}

# Saving the index for these columns
valid_columns <- which(apply(p_value_matrix, 2, function(col) all(is.na(col) | col >= 0.05)))

# Saving the index for these rows
valid_rows <- apply(p_value_matrix, 2, function(col) if(all(is.na(col) | col >= 0.05)) which.max(col) else NA)

# Create a data.frame with the Results
Result_indices <- data.frame(Column_Index = valid_columns, Row_Index = valid_rows[valid_columns])

# Verify Results
print(Result_indices)


```

```{r script_full_matrix}
  # Script for the de Wilcox group test. Now only epsilon, change such value for HV for the other indicator
  
  ## Prototype matrix for the p-values
  conf_type_aux <- unique(Problems$ID_Type_C)
  matrix_aux <- matrix(0, ncol = length(conf_type_aux), nrow = length(conf_type_aux))
  colnames(matrix_aux) <- conf_type_aux
  rownames(matrix_aux) <- conf_type_aux
  
  ## We complete the matrix for each configuration
  matrices_list <- unique(Problems$ID_Type_P) %>% lapply(function(p_type){
    
    aux <- Problems %>% filter(ID_Type_P == p_type)
    
    for(row in rownames(matrix_aux)){
      
      aux1 <- aux %>% filter(ID_Type_C == row) %>% pull(Epsilon)
      
      for(col in colnames(matrix_aux)){
        
        if(row == col){
          
          matrix_aux[row,col] <- 1
          
        } else {
          
          aux2 <- aux %>% filter(ID_Type_C == col) %>% pull(Epsilon)
          matrix_aux[row,col] <- wilcox.test(aux1, aux2, alternative = "less", exact = F)$p.value
          #less for epsilon, greater for HV
        }
      }
    }
    
    return(matrix_aux)
  })

  
```

```{r}
#Saving the matrices in files

for (i in 1:256){
  write.csv2(matrices_list[i],sprintf("files/Matrices/Epsilon/Epsilon_Exp%s.csv",i), row.names = FALSE)
}
```


```{r}
#Showing data in sunbursts (hardcoded)

library(plotly)

psos_hv <- c(246, 135, 33, 33, 28, 38)
ga_hv <- c(10,11,16)
conf_ids_hv <- c("HV", "PSO", "GA", "GOther", "G169", "G163", "P11", "P9", "P12", "P4", "P16", "POther")
conf_labels_hv <- c("HV", "PSO", "GA", "Other", "G169", "G163", "P11", "P9", "P12", "P4", "P16", "Other")
conf_parents_hv <- c("", "HV", "HV", "GA", "GA", "GA", "PSO", "PSO", "PSO", "PSO", "PSO", "PSO")

psos_t_hv <- sum(psos_hv)
ga_t_hv <- sum(ga_hv)

total_hv <- psos_t_hv + ga_t_hv

fig_hv <- plot_ly(
  
  ids = conf_ids_hv,
  labels = conf_labels_hv,
  parents = conf_parents_hv,
  values = c(total_hv, psos_t_hv, ga_t_hv, ga_hv, psos_hv),
  type = 'sunburst',
  branchvalues = 'total',
  sort = FALSE,
  insidetextorientation='radial'
)

fig_hv <- fig_hv %>% 
  layout(
    sunburstcolorway = c("#3280ae","#8cb849"),
          extendsunburstcolors = TRUE
  )
fig_hv

hvs <- c(251, 141, 33, 33, 28, 38)
conf_ids_hv <- c("HV", "P11", "P9", "P12", "P4", "P16", "POther")
conf_labels_hv <- c("HV only PSOs", "P11", "P9", "P12", "P4", "P16", "Other")
conf_parents_hv <- c("", "HV","HV","HV","HV","HV","HV")

hv <- sum(hvs)

fig_hv <- plot_ly(
  
  ids = conf_ids_hv,
  labels = conf_labels_hv,
  parents = conf_parents_hv,
  values = c(hv, hvs),
  type = 'sunburst',
  branchvalues = 'total',
  sort = FALSE,
  insidetextorientation='radial'
)

fig_hv <- fig_hv %>% 
  layout(
    sunburstcolorway = c("#3280ae","#3280ae","#3280ae","#3280ae","#3280ae","#3280ae"),
          extendsunburstcolors = TRUE
  )
fig_hv

hvs <- c(236, 210, 41)
conf_ids_hv <- c("HV","G169", "G163", "GOther")
conf_labels_hv <- c("HV only GAs", "G169", "G163", "Other")
conf_parents_hv <- c("", "HV", "HV","HV")

hv <- sum(hvs)

fig_hv <- plot_ly(
  
  ids = conf_ids_hv,
  labels = conf_labels_hv,
  parents = conf_parents_hv,
  values = c(hv, hvs),
  type = 'sunburst',
  branchvalues = 'total',
  sort = FALSE,
  insidetextorientation='radial'
)

fig_hv <- fig_hv %>% 
  layout(
    sunburstcolorway = c("#8cb849","#8cb849","#8cb849"),
          extendsunburstcolors = TRUE
  )
fig_hv





psos_eps <- c(168, 139, 105, 62, 31, 34)
ga_eps <- c(33,21,34,53)
conf_ids_eps <- c("ϵ-Indicator", "PSO", "GA", "GOther", "G163", "G166", "G118", "P11", "P9", "P3", "P1", "P15", "POther")
conf_labels_eps <- c("ϵ-Indicator", "PSO", "GA", "Other", "G163", "G166", "G118", "P11", "P9", "P3", "P1", "P15", "Other")
conf_parents_eps <- c("", "ϵ-Indicator", "ϵ-Indicator", "GA", "GA", "GA", "GA", "PSO", "PSO", "PSO", "PSO", "PSO", "PSO")

psos_t_eps <- sum(psos_eps)
ga_t_eps <- sum(ga_eps)

total_eps <- psos_t_eps + ga_t_eps

fig_eps <- plot_ly(
  
  ids = conf_ids_eps,
  labels = conf_labels_eps,
  parents = conf_parents_eps,
  values = c(total_eps, psos_t_eps, ga_t_eps, ga_eps, psos_eps),
  type = 'sunburst',
  branchvalues = 'total',
  sort = FALSE,
  insidetextorientation='radial'
)

fig_eps <- fig_eps %>% 
  layout(
    sunburstcolorway = c("#3280ae",rgb(233/255, 181/255, 69/255)),
          extendsunburstcolors = TRUE
  )
fig_eps


eps <- c(239, 214, 111, 67, 52, 52, 63)
conf_ids_eps <- c("ϵ-Indicator", "P11", "P9", "P3", "P1", "P15", "P13", "POther")
conf_labels_eps <- c("ϵ-Indicator only PSOs", "P11", "P9", "P3", "P1", "P15", "P13", "Other")
conf_parents_eps <- c("", "ϵ-Indicator", "ϵ-Indicator", "ϵ-Indicator", "ϵ-Indicator", "ϵ-Indicator", "ϵ-Indicator", "ϵ-Indicator")

ep <- sum(eps)

fig_eps <- plot_ly(
  
  ids = conf_ids_eps,
  labels = conf_labels_eps,
  parents = conf_parents_eps,
  values = c(ep, eps),
  type = 'sunburst',
  branchvalues = 'total',
  sort = FALSE,
  insidetextorientation='radial'
)

fig_eps <- fig_eps %>% 
  layout(
    sunburstcolorway = c("#3280ae","#3280ae","#3280ae","#3280ae","#3280ae","#3280ae","#3280ae"),
          extendsunburstcolors = TRUE
  )
fig_eps

eps <- c(117, 83, 67, 53, 34, 19)
conf_ids_eps <- c("ϵ-Indicator", "G199", "G169", "G163", "G118", "G166", "GOther")
conf_labels_eps <- c("ϵ-Indicator only GAs", "G199", "G169", "G163", "G118", "G166", "Other")
conf_parents_eps <- c("", "ϵ-Indicator", "ϵ-Indicator", "ϵ-Indicator", "ϵ-Indicator", "ϵ-Indicator", "ϵ-Indicator")

ep <- sum(eps)

fig_eps <- plot_ly(
  
  ids = conf_ids_eps,
  labels = conf_labels_eps,
  parents = conf_parents_eps,
  values = c(ep, eps),
  type = 'sunburst',
  branchvalues = 'total',
  sort = FALSE,
  insidetextorientation='radial'
)

fig_eps <- fig_eps %>% 
  layout(
    sunburstcolorway = c("#8cb849","#8cb849","#8cb849","#8cb849","#8cb849","#8cb849"),
          extendsunburstcolors = TRUE
  )
fig_eps


```

```{r}
#adjusting labels and positions to be prettier
library(plotly)
hvs <- c(251, 141, 33, 33, 28, 38)
conf_ids_hv <- c("HV", "P11", "P9", "P12", "P4", "P16", "POther")
conf_labels_hv <- c("HV only PSOs", "P11", "", "P12", "P4", "P16", "Other")
conf_parents_hv <- c("", "HV","HV","HV","HV","HV","HV")

hv <- sum(hvs)

fig_hv <- plot_ly(
  ids = conf_ids_hv,
  labels = conf_labels_hv,
  parents = conf_parents_hv,
  values = c(hv, hvs),
  type = 'sunburst',
  branchvalues = 'total',
  sort = FALSE,
  insidetextorientation = 'radial',
  margin = list(l = 0, r = 0, t = 0, b = 0)

)

fig_hv <- fig_hv %>% 
  layout(
    sunburstcolorway = c("#3280ae","#3280ae","#3280ae","#3280ae","#3280ae","#3280ae"),
    extendsunburstcolors = TRUE,
    annotations = list(
      list(
        x = 0.32,
        y = 0.22,
        text = "P9",
        showarrow = FALSE,
        font = list(size = 13, color = "white"),
        textangle = -45
      )
    ),
    margin = list(l = 0, r = 0, t = 0, b = 0)
  )

fig_hv
```

```{r}
psos_hv <- c(246, 135, 33, 33, 28, 38)
ga_hv <- c(10,11,16)
conf_ids_hv <- c("HV", "PSO", "GA", "GOther", "G169", "G163", "P11", "P9", "P12", "P4", "P16", "POther")
conf_labels_hv <- c("HV", "PSO", "GA", "Other", "G169", "G163", "P11", "", "P12", "P4", "P16", "Other")
conf_parents_hv <- c("", "HV", "HV", "GA", "GA", "GA", "PSO", "PSO", "PSO", "PSO", "PSO", "PSO")

psos_t_hv <- sum(psos_hv)
ga_t_hv <- sum(ga_hv)

total_hv <- psos_t_hv + ga_t_hv

fig_hv <- plot_ly(
  
  ids = conf_ids_hv,
  labels = conf_labels_hv,
  parents = conf_parents_hv,
  values = c(total_hv, psos_t_hv, ga_t_hv, ga_hv, psos_hv),
  type = 'sunburst',
  branchvalues = 'total',
  sort = FALSE,
  insidetextorientation='radial'
)

fig_hv <- fig_hv %>% 
  layout(
    sunburstcolorway = c("#3280ae","#8cb849"),
      extendsunburstcolors = TRUE,
      annotations = list(
      list(
        x = 0.265, 
        y = 0.265, 
        text = "P9", 
        showarrow = FALSE,
        font = list(size = 13, color = "white"),
        textangle = -40 
      )
    ),
    margin = list(l = 0, r = 0, t = 0, b = 0)
  )
fig_hv
```

```{r}
hvs <- c(236, 210, 41)
conf_ids_hv <- c("HV","G169", "G163", "GOther")
conf_labels_hv <- c("HV only GAs", "G169", "", "Other")
conf_parents_hv <- c("", "HV", "HV","HV")

hv <- sum(hvs)

fig_hv <- plot_ly(
  
  ids = conf_ids_hv,
  labels = conf_labels_hv,
  parents = conf_parents_hv,
  values = c(hv, hvs),
  type = 'sunburst',
  branchvalues = 'total',
  sort = FALSE,
  insidetextorientation='radial'
)

fig_hv <- fig_hv %>% 
  layout(
    sunburstcolorway = c("#8cb849","#8cb849","#8cb849"),
# amarillo: rgb(233/255, 181/255, 69/255)
          extendsunburstcolors = TRUE,
      annotations = list(
      list(
        x = 0.45,
        y = 0.1,
        text = "G163", 
        showarrow = FALSE,
        font = list(size = 12, color = "black"),
        textangle = -80 
      )
    ),
    margin = list(l = 0, r = 0, t = 0, b = 0)
  )
fig_hv
```

```{r}
psos_eps <- c(168, 139, 105, 62, 31, 34)
ga_eps <- c(33,21,34,53)
conf_ids_eps <- c("ϵ-Indicator", "PSO", "GA", "GOther", "G163", "G166", "G118", "P11", "P9", "P3", "P1", "P15", "POther")
conf_labels_eps <- c("ϵ-Indicator", "PSO", "GA", "Other", "G163", "G166", "G118", "P11", "P9", "", "P1", "P15", "Other")
conf_parents_eps <- c("", "ϵ-Indicator", "ϵ-Indicator", "GA", "GA", "GA", "GA", "PSO", "PSO", "PSO", "PSO", "PSO", "PSO")

psos_t_eps <- sum(psos_eps)
ga_t_eps <- sum(ga_eps)

total_eps <- psos_t_eps + ga_t_eps

fig_eps <- plot_ly(
  
  ids = conf_ids_eps,
  labels = conf_labels_eps,
  parents = conf_parents_eps,
  values = c(total_eps, psos_t_eps, ga_t_eps, ga_eps, psos_eps),
  type = 'sunburst',
  branchvalues = 'total',
  sort = FALSE,
  insidetextorientation='radial'
)

fig_eps <- fig_eps %>% 
  layout(
    sunburstcolorway = c("#3280ae","#8cb849"),
          extendsunburstcolors = TRUE,
      annotations = list(
      list(
        x = 0.23,
        y = 0.42,
        text = "P3", 
        showarrow = FALSE,
        font = list(size = 12, color = "white"),
        textangle = -15 
      )
    ),
    margin = list(l = 0, r = 0, t = 0, b = 0)
  )
fig_eps
```

```{r}


eps <- c(239, 214, 111, 67, 52, 52, 63)
conf_ids_eps <- c("ϵ-Indicator", "P11", "P9", "P3", "P1", "P15", "P13", "POther")
conf_labels_eps <- c("ϵ-Indicator only PSOs", "P11", "", "P3", "P1", "P15", "P13", "Other")
conf_parents_eps <- c("", "ϵ-Indicator", "ϵ-Indicator", "ϵ-Indicator", "ϵ-Indicator", "ϵ-Indicator", "ϵ-Indicator", "ϵ-Indicator")

ep <- sum(eps)

fig_eps <- plot_ly(
  
  ids = conf_ids_eps,
  labels = conf_labels_eps,
  parents = conf_parents_eps,
  values = c(ep, eps),
  type = 'sunburst',
  branchvalues = 'total',
  sort = FALSE,
  insidetextorientation='radial'
)

fig_eps <- fig_eps %>% 
  layout(
    sunburstcolorway = c("#3280ae","#3280ae","#3280ae","#3280ae","#3280ae","#3280ae","#3280ae"),
          extendsunburstcolors = TRUE,
      annotations = list(
      list(
        x = 0.27,
        y = 0.62,
        text = "P9", 
        showarrow = FALSE,
        font = list(size = 12, color = "WHITE"),
        textangle = 25 
      )
    ),
    margin = list(l = 0, r = 0, t = 0, b = 0)
  )
fig_eps
```

```{r}

eps <- c(117, 83, 67, 53, 34, 19)
conf_ids_eps <- c("ϵ-Indicator", "G199", "G169", "G163", "G118", "G166", "GOther")
conf_labels_eps <- c("ϵ-Indicator only GAs", "G199", "", "G163", "G118", "G166", "Other")
conf_parents_eps <- c("", "ϵ-Indicator", "ϵ-Indicator", "ϵ-Indicator", "ϵ-Indicator", "ϵ-Indicator", "ϵ-Indicator")

ep <- sum(eps)

fig_eps <- plot_ly(
  
  ids = conf_ids_eps,
  labels = conf_labels_eps,
  parents = conf_parents_eps,
  values = c(ep, eps),
  type = 'sunburst',
  branchvalues = 'total',
  sort = FALSE,
  insidetextorientation='radial'
)

fig_eps <- fig_eps %>% 
  layout(
    sunburstcolorway = c("#8cb849","#8cb849","#8cb849","#8cb849","#8cb849","#8cb849"),
          extendsunburstcolors = TRUE,
      annotations = list(
      list(
        x = 0.26,
        y = 0.7,
        text = "G169", 
        showarrow = FALSE,
        font = list(size = 12, color = "black"),
        textangle = 30 
      )
    ),
    margin = list(l = 0, r = 0, t = 0, b = 0)
  )
fig_eps
```